---
title: "Brendan Cullen's CV"
author: Brendan Cullen
date: "`r Sys.Date()`"
output:
  pagedown::html_resume:
    css: ['css/styles.css', 'resume']
    # set it to true for a self-contained HTML page but it'll take longer to render
    self_contained: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  results='asis', 
  echo = FALSE
)

library(glue)
library(tidyverse)

# ======================================================================
# These variables determine how the the data is loaded and how the exports are
# done.

# Is data stored in google sheets? If no data will be gather from the csvs/
# folder in project
using_googlesheets <- TRUE

# Just the copied URL from the sheet
positions_sheet_loc <- "https://docs.google.com/spreadsheets/d/1Zs3UoRaZEXoNiX-RFWNHTnr5KL8SCB_e-bik5mftt-k/"

# Is this sheet available for anyone to read? If you're using a private sheet
# set this to false and go to gather_data.R and run the data loading manually
# once to cache authentication
sheet_is_publicly_readable <- TRUE

# Is the goal of this knit to build a document that is exported to PDF? If so
# set this to true to have links turned into footnotes at the end of the
# document
PDF_EXPORT <- FALSE


CV_PDF_LOC <- "cullen_cv.pdf"
CV_HTML_LOC <- "https://bcullen.rbind.io/cv/index.html"


# A global (gasp) variable that holds all the links that were inserted for
# placement at the end
links <- c()

# ======================================================================
# Now we source two external scripts. One contains functions for building the
# text output and the other loads up our data from either googlesheets or csvs

# Functions for building sections from CSV data
source('parsing_functions.R') 

# Load data for CV/Resume
source('gather_data.R')
```



```{r}
# When in export mode the little dots are unaligned, so fix that. 
if(PDF_EXPORT){
  cat("
  <style>
  :root{
    --decorator-outer-offset-left: -6.5px;
  }
  </style>")
}
```


Aside
================================================================================

Contact {#contact}
--------------------------------------------------------------------------------

```{r}
contact_info %>% 
  glue_data("- <i class='fa fa-{icon}'></i> [{contact}]({link})")
```

Data Science Specialization {#training}
--------------------------------------------------------------------------------

As part of my doctoral training, I am currently pursuing a 5-course [Data Science Specialization](https://github.com/uo-datasci-specialization) taught entirely in R, focused on reproducible analysis, data visualization, functional programming, and machine learning within a tidyverse framework.

Programming Skills {#skills}
--------------------------------------------------------------------------------

```{r}
build_skill_bars(skills)
```


Disclaimer {#disclaimer}
--------------------------------------------------------------------------------

Created with [pagedown](https://github.com/rstudio/pagedown).
<br>Source code on [GitHub](https://github.com/brendanhcullen/bcullen/tree/master/static/cv). <i class="fa fa-github"></i>
<br>Last updated on `r Sys.Date()`.


```{r}
# When in export mode the little dots are unaligned, so fix that.
if(PDF_EXPORT){
  glue("View this CV online with links at _{CV_HTML_LOC}_")
} else {
  glue("[<i class='fas fa-download'></i> Download a PDF of this CV]({CV_PDF_LOC})")
}
```


Main
================================================================================

Brendan Cullen {#title}
--------------------------------------------------------------------------------

```{r}
print_text_block(text_blocks, 'intro')
```


Education {data-icon=graduation-cap data-concise=true}
--------------------------------------------------------------------------------

```{r}
print_section(position_data, 'education')
```

Certification {data-icon=certificate data-concise=true}
--------------------------------------------------------------------------------


```{r}
print_section(position_data, 'certification')
```


Honors & Awards {data-icon=trophy}
--------------------------------------------------------------------------------

```{r}
print_section(position_data, 'honors_awards')
```


Professional Research Experience {data-icon=suitcase}
--------------------------------------------------------------------------------

```{r}
print_section(position_data, 'research_positions')
```


Teaching Experience (Courses) {data-icon=chalkboard-teacher}
--------------------------------------------------------------------------------

::: aside
```{r}
print_text_block(text_blocks, 'teaching_experience_aside')
```
:::

```{r}
print_section(position_data, 'teaching_positions')
```

Workshops Taught {data-icon=laptop}
--------------------------------------------------------------------------------

```{r}
print_section(position_data, 'workshops')
```

<br><br>

Peer-reviewed Publications{data-icon=book}
--------------------------------------------------------------------------------

::: aside
```{r}
print_text_block(text_blocks, 'publications_aside')
```
:::

```{r}
print_section(position_data, 'academic_articles')
```

Pre-prints & Works in Progress{data-icon=edit}
--------------------------------------------------------------------------------

```{r}
print_section(position_data, 'works_in_progress')
```


Selected Poster Presentations{data-icon=image}
--------------------------------------------------------------------------------

```{r}
print_section(position_data, 'posters')
```

<br>

Talks{data-icon=users}
--------------------------------------------------------------------------------

```{r}
print_section(position_data, 'talks')
```

<br><br><br><br>

Additional Training {data-icon=user-graduate}
--------------------------------------------------------------------------------


```{r}
print_section(position_data, 'training')
```

Service & Leadership{data-icon=hands-helping}
--------------------------------------------------------------------------------


```{r}
print_section(position_data, 'service')
```


```{r}
if(PDF_EXPORT){
  cat("
  
Links {data-icon=link}
--------------------------------------------------------------------------------

<br>


")
  
  walk2(links, 1:length(links), function(link, index){
    print(glue('{index}. {link}'))
  })
}
```


