---
title: Brendan Cullen's Résumé
author: Brendan Cullen
date: "`r Sys.Date()`"
output:
  pagedown::html_resume:
    css: ['css/custom_resume.css', 'css/styles.css', 'resume']
    # set it to true for a self-contained HTML page but it'll take longer to render
    self_contained: true
---


```{r, include=FALSE}
knitr::opts_chunk$set(
  results='asis', 
  echo = FALSE
)
library(tidyverse)
library(glue)

# ======================================================================
# These variables determine how the the data is loaded and how the exports are
# done.

# Is data stored in google sheets? If no data will be gather from the csvs/
# folder in project
using_googlesheets <- TRUE

# Just the copied URL from the sheet
positions_sheet_loc <- "https://docs.google.com/spreadsheets/d/1Zs3UoRaZEXoNiX-RFWNHTnr5KL8SCB_e-bik5mftt-k/"

# Is this sheet available for anyone to read? If you're using a private sheet
# set this to false and go to gather_data.R and run the data loading manually
# once to cache authentication
sheet_is_publicly_readable <- TRUE

# Is the goal of this knit to build a document that is exported to PDF? If so
# set this to true to have links turned into footnotes at the end of the
# document
PDF_EXPORT <- FALSE


# A global (gasp) variable that holds all the links that were inserted for
# placement at the end
links <- c()

# ======================================================================
# Now we source two external scripts. One contains functions for building the
# text output and the other loads up our data from either googlesheets or csvs

# Functions for building sections from CSV data
source('parsing_functions.R') 

# Load data for CV/Resume
source('gather_data.R')

# Now we just need to filter down the position data to include less verbose
# categories and only the entries we have designated for the resume
position_data <- position_data %>% 
  filter(in_resume) %>% 
  mutate(
    # Build some custom sections by collapsing others
    section = case_when(
      section %in% c('research_positions', 'teaching_positions') ~ 'positions', 
      section %in% c('academic_articles') ~ 'writings',
      #section %in% c('workshops') ~ 'workshops',
      TRUE ~ section
    )
  ) 
```



Aside
================================================================================


Contact <i class="fa fa-address-book"></i> {#contact}
--------------------------------------------------------------------------------

```{r}
contact_info %>% 
  glue_data("- <i class='fa fa-{icon}'></i> [{contact}]({link})")
```

<br>

Certification <i class="fab fa-r-project"></i> {#certification}
--------------------------------------------------------------------------------

I am an [RStudio Certified Instructor](https://education.rstudio.com/trainers/people/cullen+brendan/) in the [tidyverse](https://www.tidyverse.org/) (see [certificate](https://github.com/brendanhcullen/rstudio-instructor-certification/blob/master/certificate.png)).

<br>

Other Training <i class="fa fa-user-graduate"></i> {#training}
--------------------------------------------------------------------------------

As part of my doctoral training, I am currently pursuing a 5-course [Data Science Specialization](https://github.com/uo-datasci-specialization) taught entirely in R, focused on reproducible analysis, data visualization, functional programming, and machine learning within a tidyverse framework.

<br>

Programming Skills <i class="fa fa-terminal"></i> {#skills}
--------------------------------------------------------------------------------

```{r}
build_skill_bars(skills)
```

<br>

Philosophy <i class="fa fa-lightbulb"></i> {#philosophy}
--------------------------------------------------------------------------------

I am a strong advocate for open science and open-source data science software. I particularly enjoy helping others learn data science skills, and I value kindness and inclusivity above all else in my teaching.

<br>

More info <i class="fa fa-info-circle"></i> {#more-info}
--------------------------------------------------------------------------------

See full CV at [bcullen.rbind.io/cv](https://bcullen.rbind.io/cv/index.html) for more complete list of positions and [peer-reviewed publications](https://scholar.google.com/citations?user=02RYe4UAAAAJ&hl=en).

<br> 

Disclaimer {#disclaimer}
--------------------------------------------------------------------------------

Created with [pagedown](https://github.com/rstudio/pagedown).
<br>Source code on [GitHub](https://github.com/brendanhcullen/bcullen/tree/master/static/cv). <i class="fa fa-github"></i>
<br>Last updated on `r Sys.Date()`.


```{r}
RESUME_PDF_LOC <- "cullen_resume.pdf"
glue("[<i class='fas fa-download'></i> Download a PDF copy]({RESUME_PDF_LOC})")
```

Main
================================================================================

Brendan Cullen {#title}
--------------------------------------------------------------------------------

```{r}
print_text_block(text_blocks, 'intro')
```



Education {data-icon=graduation-cap data-concise=true}
--------------------------------------------------------------------------------

```{r}
position_data %>% print_section('education_resume')
```



Selected Positions {data-icon=suitcase}
--------------------------------------------------------------------------------

```{r}
position_data %>% print_section('positions_resume')
```


Selected Workshops Taught {data-icon=chalkboard-teacher}
--------------------------------------------------------------------------------

```{r}
position_data %>% print_section('workshops_resume')
```


Selected Service {data-icon=hands-helping}
--------------------------------------------------------------------------------

```{r}
position_data %>% print_section('service')
```

