---
title: Teaching an R Bootcamp Remotely
author: Brendan Cullen
date: '2020-10-19'
slug: []
categories: []
tags:
  - data science
  - education
  - R
subtitle: ''
summary: 'Lessons learned from my first experience teaching R remotely and a collection of resources & tools I found really useful'
authors: []
lastmod: '2020-10-19T10:26:32-07:00'
featured: no
image:
  caption: '<span>Photo by <a href="https://unsplash.com/@martinadams?utm_source=unsplash&amp;utm_medium=referral&amp;utm_content=creditCopyText">Martin Adams</a> on <a href="https://unsplash.com/s/photos/sky?utm_source=unsplash&amp;utm_medium=referral&amp;utm_content=creditCopyText">Unsplash</a></span>'
  focal_point: ''
  preview_only: no
projects: []
disable_codefolding: false
codefolding_show: "hide"
codefolding_nobutton: true 
---

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(googlesheets4)
library(patchwork)
library(Manu)
```


# Background

Back in Fall 2018, the Department of Psychology at the University of Oregon started to offer an introductory R Bootcamp the week before the start of the new academic year. It was my privilege to be able to lead the 3rd Annual UO Psych Bootcamp this year. 


# Who attended? 

```{r, warning=FALSE, message=FALSE}
# suspend authorization so no need for a token for accessing Google Drive data
googlesheets4::gs4_deauth()

# import bootcamp registration data 
registration <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1wcGdQU7Vkg3XLvu0mArQx_gcSt0xHztbXmAmnj_tqmE/edit?usp=sharing")

# clean data
plot_data <- registration %>% 
  janitor::clean_names() %>% 
  select(position = what_is_your_position_at_uo,
         experience = do_you_have_any_experience_in_r_no_experience_is_required) %>%
  mutate(position = factor(case_when(position %in% c("Undergraduate", "Undergraduate student") ~ "Undergrad Student", 
                                     position %in% c("RA", "Lab Manager", "Data Analyst") ~ "Research Staff",
                              TRUE ~ as.character(position))),
         experience = factor(experience, levels = c(
           "I've never heard of it",
           "I've heard of it but have never used it",
           "I've opened it before",
           "I've used it to analyze data",
           "I use it often to analyze data")))

# plot positions at UO
theme_set(theme_minimal(base_size = 15)) 
theme_update(legend.position = "none", 
          plot.title.position = "plot",
          panel.grid.minor.y = element_blank())

plot_data %>% 
  count(position) %>% 
  ggplot(aes(fct_reorder(position, n), n, fill = position)) + 
  geom_col() +
  coord_flip() + 
  scale_fill_manual(values = get_pal("Kotare")) + 
  labs(x = "", y = "Count", title = "Participants' roles at University of Oregon")
```


Prior experience

```{r}
# plot prior experience 
plot_data %>% 
  count(experience) %>% 
  ggplot(aes(experience, n, fill = experience)) + 
  geom_col() +
  coord_flip() + 
  scale_fill_manual(values = get_pal("Kotare"))  + 
  labs(x = "", y = "Count", title = "Participants' prior experience with R")
```


# Logistics 

# Resources & Recommendations

## Tools I learned 

## Resources I used

# Feedback 

